### 2023-04-27 chapters/analysis/A-CdM.tex
Condition monitoring complete before proofreading

The defect needs to be either repaired or replaced, preferably without significant production downtime, further damage to the other attached elements, or any endangerment of the responsible personnel. The maintenance strategies are chosen according to the machine's importance as a result of its failure effect evaluation on the system. The guide to set appropriate maintenance procedures is outlined by IEC 60706-2 standard and involves reliability-centered maintenance (RCM) analysis \cite{el-thalji_predictive_2019}.

@@

\textbf{Reactive maintenance} allows machinery to run to complete failure. This is the most inappropriate way to maintain the production line, but it is straightforward. It requires a large stock of replacement parts on-site and breakage inflict a `crisis management mode' upon the plant \cite{scheffer_practical_2004}. On-demand repairs are justified for consumer products or in the factory capable of fully and quickly replacing the halted machine with a backup. 
\bigbreak

\textbf{Preventive maintenance} is performed before any issue is detected. Maintenance occurs at regular intervals derived from a predetermined period in the calendar or expected machine running time (e.g. MTTF - Mean Time To Failure). The schedule is crucial but can result in components being replaced in good condition when further utilization is possible or too late after the machine breaks. In this case, conservative planning is usually the norm to keep machine always in perferct state and therefore more frequent intervention. \cite{mohanty_machinery_2015}.  
\bigbreak

\textbf{Predictive maintenance} known as condition-based maintanance (CbM), improves the predictability of reactive maintenance and eliminates the waste in overall resource utilization of cautious prevention. The machine downtime is scheduled after the detection of unhealthy trends in fault monitoring with sensors and troublesome components are identified.

Measurable decrease in efectivity allows us to order necessary parts in advance and organize repairs of several machines at a convenient time. The misdetection leads to increased costs compared to previous methods and raises the expectation that faults are distinguishable among themselves~\cite{davies_handbook_2012}.

\bigbreak

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{assets/P-F-Curve.png}
	\caption{P-F curve represets evolution of the asset's health \cite{jennions_integrated_2011}}
	\label{fig:p-f-curve}
\end{figure}

@@

High failure rate is present not only at the worn out stage when the parts are fatigued or corroded, but also in the early stages soon after assembly. Causes can be found in the manufacturing or material defects, inadequate installation, or improper start-up procedures. During the stable middle phase the malfunction can occur after machine excessive overload. The time plot to failure rate is known as the bath tub curve (Fig.~\ref{fig:bath-tub-curve}).

\subsection{Vibration fault types}
Mechanical problems during machinery operation bring about vibrations in the vast majority of cases. The cause of vibration comes out of the changing force in its magnitude or direction. The most emerging defects can be encompassed by explaining the deficiencies of the mechanical structure broadly categorized as \textbf{unbalance, misalignment, looseness, excentricity, and influence of the external force}~\cite{davies_handbook_2012}.

However, rotating machine disorders do exhibit frequency signatures at various ranges in the frequency spectrum. Imbalance, misalignment, and looseness normally appear at frequencies up to 300 Hz. These low-frequency faults are associated with the movement of the shaft and primarily coincide with revolution speed and its harmonics. Bearing and gearbox defects in the late stages of development, show up in the range between 300 Hz to 1 kHz. Higher frequencies measured traditionally to a limit of 10 kHz help notice the flaw of bearings even sooner~\cite{torres_automatic_2022}.

\begin{itemize}
\item predominant frequency
\item synchronous frequency
\item subsynchronous frequency
\item fundamental frequency
\item harmonic frequency
\item subharmonic frequency
\end{itemize}
% p.294 Davies

\cite{scheffer_practical_2004} p.98 - 141

\cite{davies_handbook_2012} p.282
% Automatic Anomaly Detection in Vibration Analysis Based on Machine Learning Algorithms
\cite{torres_automatic_2022}
% Vibration Guide
\cite{noauthor_vibration_2000}
% The experimental application of popular machine learning algorithms on predictive maintenance and the design of IIoT based condition monitoring system
\cite{cakir_experimental_2021}
% Technická diagnostika
\cite{ziaran_technicka_2013}

%TODO
Why monitor with vibrations,
- Resonance frequencies of each part - machine must run at speeds not aligned with resonance frequencies - Campbell diagram - task for mechanical engineers
- Faults - reasons and frequency content

% Describe by frequency responses and assign them possibale faults instead

\begin{itemize}
\itemsep0pt
\item Synchrounous response - based on RPM
\item Mass unbalance
\item Misalignment
\item Eccentricity
\item Bent or bow shaft
\item Cracked shaft
\item Rotor rubs - friction
\item Looseness
\item Auxiliery mechanical systems: Gearbox, Bearings, Belt 
\end{itemize}

% Bandsaws
% Vibration of bandsaws
\cite{lengoc_vibration_1990}
% Study on Online Detection and Fault Diagnosis of Band Saw Equipment
\cite{chen_study_2014}

@@

Criteria introduced to judge vibration severity are its absolute vibration magnitude, change in the magnitude vector, and rate of change. In terms of maximal magnitudes the machines of varied sizes are split into four severity zones defined in the chart ~(Tab.~\ref{tab:iso20816-vibration-severity}). The table values serve as guidelines towards realistic requirements between machinemanufacturers and customers.

@@

\caption{ISO 20816 vibration serverity chart \cite{noauthor_iso_2016}}

@@

\emph{Zone A} is reserved for newly commissioned machines. \emph{Zone B} signifies suitability for long-term operation. In \emph{zone C} is the machine deemed in unsatisfactory condition and corrective action should be taken soon. Finally, in zone D vibrations can cause damage to the machine. The span of acceptable values differs on machine class from I through to IV with an output power of 15 kW (class I), 75 kW (class II), 10 MW (class III), or greater.

The operational limits in the form of \emph{alarms} and \emph{trips} are usually established on the zone boundaries or close to them. Alarms are placed between zones B and C and provide a warning about reaching the threshold significant for noticeable change. Trips in between zones C and D urge immediate action or machine shut down.

@@

@@

The transducer of choice for condition monitoring is the accelerometer which can provide the acceleration value of the body and velocity after signal integration. However, standard cautions against double integrating for displacement. The recommended frequency range for an accelerometer is 0.1 Hz to 30 kHz. The choice of transducer mount significantly lowers its resonance frequency which is least influenced by stud mount and stiff cement mount. The resonance is reduced to around 8 kHz with the use of soft epoxy or permanent magnet.

@@

Broadband measurement requires `frequency ranges of 0.2 times the lowest rotational frequency to the highest frequency of interest' \cite{noauthor_iso_2002}, not exceeding 10 kHz, with RMS velocity 0.1 - 100 mm/s. Bearings and gears diagnosis may push the upper-frequency limit even higher. The tolerances of amplitude and frequency calibrations fall into two types with allowable tolerances of $\pm 5 \%$ or $\pm 10 \%$ (Fig.~\ref{fig:tranducer-response}).

% Different under load 
% Baseline measurement - what are those p.28
% Monitoring programme p.12 (vibrations differ under load)
% Later: Potencial causes for faults (p. 45) - use in vibration fault types

\begin{enumerate}
\itemsep0pt
\item Review machinery history and establish failure modes.
\item When vibration monitoring is not applicable check for other condition monitoring techniques or use preventive maintance routine.
\item Select monitoring points
\item Take preliminary vibration measurements
\item Select vibration monitoring techniques: broadband, frequency analysis or special techniques. Set parameters of measurements units.
\item Take baseline measurements
\item Change levels that would warrant investigation
\item Carry out routine condition monitoring
\item If alarm was exceeded notify approriate personnel, review data and trends, perform diagnostic evaluation, repair as necessary. In case new baseline is needed continue in step of taking baseline measurements.
\item Shut down machine when trip level is exceeded. Than proceed same as after alarm trigger.
\end{enumerate}

\newpage

>>>>

The defect needs to be either repaired or replaced, preferably without significant production downtime, further damage to the other attached elements, or any endangerment of the responsible personnel. The maintenance strategies are chosen according to the machine's importance as a result of its failure effect evaluation on the system. The guide to set appropriate maintenance procedures is outlined by the IEC~60706-2 standard and involves reliability-centered maintenance analysis~\cite{el-thalji_predictive_2019}.

@@

\textbf{Reactive maintenance} allows machinery to run to complete failure. This is the most inappropriate way to maintain the production line, but it is straightforward. It requires a large stock of replacement parts on-site and breakage inflicts a `crisis management mode' upon the plant \cite{scheffer_practical_2004}. On-demand repairs are justified if short downtime is an acceptable,  full, and quick replacement of a broken machine with a backup is possible or there is no threat of failure to the surrounding environment~\cite{ziaran_technicka_2013}.
\bigbreak

\textbf{Preventive maintenance} is performed before any issue is detected. Maintenance occurs at regular intervals derived from a predetermined period in the calendar or expected machine running time (e.g. MTTF - Mean Time To Failure). The schedule is crucial but can result in components being replaced in good condition when further utilization is possible or too late after the machine breaks. In this case, conservative planning is usually the norm to keep the machine always in a perfect state and therefore more frequent intervention. \cite{mohanty_machinery_2015}.
\bigbreak

\textbf{Predictive maintenance} known as condition-based maintenance (CbM), improves the predictability of reactive maintenance and eliminates the waste in overall resource utilization of cautious prevention. The machine downtime is scheduled after the detection of unhealthy trends in fault monitoring with sensors and troublesome components are identified.

A measurable decrease in effectivity allows us to order necessary parts in advance and organize repairs of several machines at a convenient time. The misdetection leads to increased costs compared to previous methods and raises the expectation that faults are distinguishable among themselves~\cite{davies_handbook_2012}.
\bigbreak

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{assets/P-F-Curve.png}
	\caption{P-F curve represents the evolution of the asset's health \cite{jennions_integrated_2011}}
	\label{fig:p-f-curve}
\end{figure}

@@

A high failure rate is present not only at the worn-out stage when the parts are fatigued or corroded but also in the early stages soon after assembly. Causes can be found in manufacturing or material defects, inadequate installation, or improper start-up procedures. During the stable middle phase, malfunction can occur after the machine's excessive overload. The time plot to failure rate is known as the bath tub curve~(Fig.~\ref{fig:bath-tub-curve}).

\subsection{Vibration fault types}
Mechanical problems during machinery operation bring about vibrations in the vast majority of cases. Therefore vibroacoustic diagnostics is considered one of the most important methods in early component fault identification~\cite{ziaran_technicka_2013}.

The cause of vibration comes out of the changing force in its magnitude or direction. The most emerging defects can be encompassed by explaining the deficiencies of the mechanical structure broadly categorized as \textbf{unbalance, misalignment, looseness, excentricity, deformation, crack and influence of the external force} (e.g. friction) ~\cite{davies_handbook_2012}. It is important to stress that our concern are not the underlying deficiencies in mechanical parts, but the correct fault classification based on the signal waveform.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{assets/complex-vibrations.png}
	\caption{Complex machinery vibrations~\cite{davies_handbook_2012}}
	\label{fig:machinery-vibrations}
\end{figure}

Rotating machine disorders do exhibit frequency signatures at various ranges in the frequency spectrum with supplementary symptoms carried in phase. Most of the occurring faults can be tied to the main rotational speed of the component under investigation (Fig.~\ref{fig:machinery-vibrations})~\cite{davies_handbook_2012}. Imbalance, misalignment, and looseness normally appear at frequencies up to 300 Hz. These low-frequency faults are associated with the movement of the shaft and primarily coincide with revolution speed and its harmonics. Bearing and gearbox defects in the late stages of development, show up in the range between 300 Hz to 1 kHz. Higher frequencies measured traditionally to a limit of 10 kHz help notice the deficiencies in bearings even sooner~\cite{torres_automatic_2022}.

One of the methods vibration experts utilize in the identification of the damaged part from the frequency spectrum is \textbf{order analysis}. In essence, it is about noticing the excessive peaks at harmonic frequencies that are integer multiples of fundamental frequency (1x RPM)~(Tab.~\ref{tab:vibration-causes}):

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|ll|l|l|}
\hline
\multicolumn{2}{|l|}{\textbf{Frequency content}}                            & \textbf{Common case}                                                                                                     & \textbf{Other causes}                                                                                                                          \\ \hline
\multicolumn{1}{|l|}{\multirow{4}{*}{Synchronous}} & 1 x RPM                & Imbalance                                                                                                                & \begin{tabular}[c]{@{}l@{}}Eccentric journals\\ Bent shaft / Misalignment\\    (high axial vibration)\\ Bad belt (if RPM of belt)\end{tabular} \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & 2 x RPM                & Looseness                                                                                                                & \begin{tabular}[c]{@{}l@{}}Misalignment \\    (high axial vibration)\\ Cracked rotor\\ Bad belt (if RPM of belt)\end{tabular}                  \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & 3 x RPM                & Misalignment                                                                                                             & and axial looseness                                                                                                                             \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & Many x RPM             & \begin{tabular}[c]{@{}l@{}}Bad gears\\ Severe looseness\end{tabular}                                                     & \begin{tabular}[c]{@{}l@{}}Gear teeth x RPM\\ Fan blade count x RPM\end{tabular}                                                               \\ \hline
\multicolumn{1}{|l|}{Subsynchrous}                 & \textless 1 x RPM      & Oil whirl                                                                                                                & \begin{tabular}[c]{@{}l@{}}Bad drive belt\\ Background\\ Resonance\end{tabular}                                                                \\ \hline
\multicolumn{1}{|l|}{Non-synchronous}              & \multicolumn{1}{c|}{-} & \begin{tabular}[c]{@{}l@{}}Electrical problems (x 50 Hz)\\ Reciprocating forces\\ Aerodynamic forces \\ Bad antifriction bearings\end{tabular} & Rubbing \\ \hline
\end{tabular}
\end{adjustbox}
\caption{Expert observed likely vibration causes (based on~\cite{davies_handbook_2012,ziaran_technicka_2013,noauthor_iso_2002})}
\label{tab:vibration-causes}
\end{table}

Because of inherent tolerances in machine manufacturing and assembly, the rotational frequency always manifests itself, even in baseline signature~\cite{davies_handbook_2012, noauthor_iso_2002}. In the most likely scenario, some faults appear as compared to rotational frequency solely in\textbf{synchronous, subsynchronous, or non-synchronous components}. The defects can occur also in a predictable combination of the ones mentioned. Other common patterns experts look for are modulation sidebands typical for bearings and gears extractable with cepstrum analysis~\cite{ziaran_technicka_2013}. Therefore procedure relying on elimination narrows down unrelated causes effectively.

@@

Criteria introduced to judge vibration severity are its absolute vibration magnitude, change in the magnitude vector, and rate of change. In terms of maximal magnitudes the machines of varied sizes are split into four severity zones defined in the chart ~(Tab.~\ref{tab:iso20816-vibration-severity}). The table values serve as guidelines towards realistic requirements between machine manufacturers and customers.

@@

\caption{ISO 20816 vibration severity chart with typical magnitudes \cite{noauthor_iso_2016}}

@@

\emph{Zone A} is reserved for newly commissioned machines. \emph{Zone B} signifies suitability for long-term operation. In \emph{zone C} is the machine deemed in unsatisfactory condition and corrective action should be taken soon. Finally, in zone D vibrations can cause damage to the machine. The span of acceptable values differs on machine class from I through to IV with an output power of 15~kW (class~I), 75~kW (class~II), 10~MW (class~III), or greater.

The operational limits in the form of \emph{alarms} and \emph{trips} are usually established on the zone boundaries or close to them. Alarms are placed between zones B and C and provide a warning about reaching the threshold significant for noticeable change. Trips in between zones C and D urge immediate action or machine shut down. Both limits should not exceed 1.25 times the upper boundary or lower zones and initially are set based on previous experience with the machine~\cite{noauthor_iso_2002}.

@@

The condition monitoring programme is run according to a flowchart adapted from one designed in the standard specifically to best benefit the plant. Those steps can be summarized as follows ~\cite{noauthor_iso_2002}:
\begin{enumerate}
\itemsep0pt
\item Review machinery history and establish failure modes.
\item When vibration monitoring is not applicable check for other condition monitoring techniques or resort to preventive maintenance.
\item Select monitoring points and take preliminary vibration measurements
\item Select vibration monitoring techniques: broadband, frequency analysis, or special techniques, and set parameters of measurement units.
\item Take baseline measurements.
\item Change levels that would warrant investigation.
\item Carry out routine condition monitoring.
\item If an alarm was exceeded, notify appropriate personnel to review data and trends, perform diagnostic evaluation, and repair as necessary. In case a new baseline is needed continue in the step of taking baseline measurements.
\item Shut down the machine when the trip level is exceeded. Then proceed the same as after the alarm trigger.
\end{enumerate}

@@

The transducer of choice for condition monitoring is the accelerometer which can provide the acceleration value of the body and velocity after signal integration. However, standard cautions against double integrating for displacement. The recommended frequency range for an accelerometer is 0.1~Hz to 30~kHz. The choice of transducer mount significantly lowers its resonance frequency which is least influenced by stud mount and stiff cement mount. The resonance is reduced to around 8~kHz with the use of soft epoxy or permanent magnet.

@@

Broadband measurement requires `frequency ranges of 0.2 times the lowest rotational frequency to the highest frequency of interest'~\cite{noauthor_iso_2002}, not exceeding 10 kHz, with RMS velocity 0.1 - 100~mm/s. Bearings and gears diagnosis may push the upper-frequency limit even higher. The tolerances of amplitude and frequency calibrations fall into two types with allowable tolerances of $\pm 5 \%$ or $\pm 10 \%$~(Fig.~\ref{fig:tranducer-response}).

Equipment's `health' can be mischaracterized when there are significant differences in the machine's normal operating conditions. Baseline measurements in all acceptable conditions are to be acquired to reduce the error in vibration evaluation. According to bath tub curve~(Fig.~\ref{fig:bath-tub-curve}) reference signatures should be obtained after the initial part wear-in period. The reference spectral mask of the baseline condition is designed if maximal acceptance amplitudes are different for each significant frequency band ~\cite{ziaran_technicka_2013}.

The vibration baseline is defined by broad-band magnitudes and phases of motion vectors, waveforms in time and frequency domain, the rotational speed of the machine as well as its frequency response to different speeds during start-up and coast-down captured in the Bode plot and waterfall plot. Changes during the machine operation are then depicted in value trends. Trends can be shown of overall amplitudes or limited to frequency bands.
